\chapter{Systematic Literature Review}
\label{ch:slr}

\section{Introduction}
\label{sec:slr-intro}

This chapter presents a systematic literature review (SLR) conducted to map
the state of the art at the intersection of process mining, stochastic
modeling, and machine learning applied to software process forecasting. The
review follows the guidelines established by Kitchenham and
Charters~\cite{kitchenham2007} for evidence-based software engineering,
complemented by the snowballing procedures described by
Wohlin~\cite{Wohlin2024} and the mapping study guidelines of Petersen et
al.~\cite{petersen2015}. The primary objective is to identify, classify, and
synthesize existing approaches that combine process-aware techniques with
predictive modeling for software development outcomes, establishing the
research gap that motivates PATHCAST proposed in
Chapter~\ref{ch:method}.

The review addresses three overarching research questions, each decomposed
into two sub-questions covering the application landscape of process mining in
SDLC contexts, the techniques and prediction methods employed, and the degree
of integration between process mining and stochastic forecasting methods.


\section{Review Protocol}
\label{sec:slr-protocol}

\subsection{Research Questions}
\label{sec:slr-rqs}

The SLR is guided by the following research questions, structured to
progressively map the landscape from application scope to technique
integration:

\begin{table}[htbp]
\centering
\caption{Research questions and expected outputs.}
\label{tab:slr-rqs}
\small
\setlength{\tabcolsep}{4pt}
\begin{tabular}{c c p{7.1cm} p{3.8cm}}
\toprule
\textbf{RQ} & \textbf{Sub-RQ} & \textbf{Question} & \textbf{Expected Output} \\
\midrule
RQ1 & RQ1.1 & Which SDLC phases have been analyzed with process mining? &
  Frequency distribution by phase \\
RQ1 & RQ1.2 & What event log sources are used and how are they constructed? &
  Taxonomy of sources and construction methods \\
\midrule
RQ2 & RQ2.1 & Which process discovery algorithms have been applied to SDLC? &
  Comparative matrix \\
RQ2 & RQ2.2 & Which predictive and stochastic techniques have been used? &
  Technique classification \\
\midrule
RQ3 & RQ3.1 & What is the level of integration between PM and stochastic
  methods? & Integration maturity assessment \\
RQ3 & RQ3.2 & What gaps exist for process-aware forecasting in the SDLC? &
  Gap analysis matrix \\
\bottomrule
\end{tabular}
\end{table}


\subsection{Search Strategy}
\label{sec:slr-search}

The search strategy combines automated database search with backward and
forward snowballing. The automated search targets five digital libraries:
IEEE~Xplore, ACM Digital Library, Scopus, Web of Science, and SpringerLink.

\subsubsection{PICO Decomposition}

The search string construction follows the PICO framework adapted for
software engineering, as shown in Table~\ref{tab:pico}.

\begin{table}[htbp]
\centering
\caption{PICO decomposition for search string construction.}
\label{tab:pico}
\begin{tabular}{lp{4cm}p{7cm}}
\toprule
\textbf{Facet} & \textbf{Role} & \textbf{Concepts} \\
\midrule
Population    & Application domain    & Software engineering, SDLC, software process,
                                        DevOps, CI/CD \\
Intervention  & Analytical technique  & Process mining, Markov chain, Monte Carlo,
                                        predictive PM, ML \\
Comparison    & N/A                   & (omitted) \\
Outcome       & Result                & Forecasting, prediction, lead time, cycle time,
                                        process model \\
\bottomrule
\end{tabular}
\end{table}

\subsubsection{Search Strings}

Three complementary search strings are executed to maximize coverage while
maintaining precision.

\paragraph{Principal String (P $\wedge$ I).}
Block~P captures the software engineering context:

\begin{quote}
\small
\texttt{("software engineering" OR "software development" OR "software process"
OR "software lifecycle" OR "software life cycle" OR "SDLC" OR "development
lifecycle" OR "development life cycle" OR "DevOps" OR "CI/CD" OR "continuous
integration" OR "continuous delivery" OR "agile development" OR "agile process"
OR "issue tracking" OR "bug tracking" OR "version control" OR "code review" OR
"pull request" OR "software project" OR "software repository" OR "software
maintenance" OR "software evolution")}
\end{quote}

Block~I captures the process mining and stochastic modeling techniques:

\begin{quote}
\small
\texttt{("process mining" OR "process discovery" OR "conformance checking" OR
"workflow mining" OR "predictive process monitoring" OR "process prediction" OR
"remaining time prediction" OR "process forecasting" OR "event log" OR "process
simulation" OR "Monte Carlo simulation" OR "Markov chain" OR "Markov model" OR
"stochastic process model" OR "transition probability" OR "absorbing chain" OR
"lead time prediction" OR "cycle time prediction")}
\end{quote}

\paragraph{Complementary String (MSR $\rightarrow$ Process Modeling).}
Captures studies using MSR terminology without canonical process mining terms:

\begin{quote}
\small
\texttt{("mining software repositories" OR "software repository mining" OR
"GitHub mining" OR "Jira mining") AND ("process model" OR "Markov" OR "Monte
Carlo" OR "stochastic" OR "lead time" OR "cycle time" OR "throughput" OR
"transition matrix")}
\end{quote}

\paragraph{Frontier String (Stochastic $\wedge$ PM $\wedge$ SE).}
Targets the intersection of stochastic methods with process mining in software
contexts:

\begin{quote}
\small
\texttt{("Monte Carlo" OR "Markov chain" OR "stochastic simulation" OR
"probabilistic forecast" OR "transition matrix") AND ("process mining" OR
"event log" OR "workflow mining" OR "process model" OR "business process") AND
("software development" OR "software engineering" OR "software process" OR
"DevOps" OR "SDLC" OR "issue tracker" OR "commit log" OR "build pipeline")}
\end{quote}

The strings are adapted to each database's syntax and field conventions as
summarized in Table~\ref{tab:db-adaptations}. The search scope covers
publications from January 1994 to December 2026.

\begin{table}[htbp]
\centering
\caption{Database-specific search adaptations and raw retrieval counts.}
\label{tab:db-adaptations}
\small
\setlength{\tabcolsep}{4pt}
\begin{tabular}{p{2.6cm} p{2.6cm} c p{5.0cm} r}
\toprule
\textbf{Database} & \textbf{Search Field} & \textbf{Strings} & \textbf{Filters} & \textbf{Records} \\
\midrule
Scopus         & TITLE-ABS-KEY & 3         & ar, cp; EN; 1994--2026                 & 2{,}377 \\
IEEE Xplore    & All Metadata  & 4 (split) & Conf.\ + Journals; 1994--2026          & 1{,}386 \\
Web of Science & TS (Topic)    & 3         & Article, Proc.; EN; 1994--2026         & 997 \\
SpringerLink   & Simplified    & 8 queries & Chapter, Article; EN                   & 2{,}884 \\
ACM DL         & Abstract      & 2         & Date: 1994--2026                       & 96 \\
Snowballing    & --            & --        & Backward + forward                     & 595 \\
Control set    & --            & --        & Manual                                 & 12 \\
\midrule
\textbf{Total} &  &  &  & \textbf{8{,}347} \\
\bottomrule
\end{tabular}
\end{table}

\subsubsection{Snowballing Procedure}

Backward snowballing examines the reference lists of all included primary
studies. Forward snowballing uses Google Scholar citation tracking to identify
subsequent works that cite the included studies. Both rounds apply the same
inclusion and exclusion criteria defined in Section~\ref{sec:slr-criteria}.

\subsubsection{Search String Validation}

The search strings are validated against a control set of 10 papers that are
known to be relevant to the review scope
(Table~\ref{tab:validation-set}). The acceptance threshold is $\geq 8/10$
papers found; the validation returned $10/10$ ($100\%$, PASS).

\begin{table}[htbp]
\centering
\caption{Control papers for search string validation.}
\label{tab:validation-set}
\small
\setlength{\tabcolsep}{4pt}
\begin{tabular}{c p{4.5cm} p{4.1cm} p{1.6cm}}
\toprule
\textbf{\#} & \textbf{Paper} & \textbf{Justification} & \textbf{Captured via} \\
\midrule
V1  & Cook \& Wolf (1998)            & Seminal PM + SE                    & ACM \\
V2  & Rubin et al. (2007)            & PM framework for SW processes      & Scopus \\
V3  & Poncin et al. (2011)           & PM + MSR intersection              & IEEE \\
V4  & Caldeira \& Cardoso (2019)     & PM + SW process assessment         & IEEE \\
V5  & Lemos et al. (2011)            & PM for SW process management       & IEEE \\
V6  & Whittaker \& Thomason (1994)   & Markov + SW testing (seminal)      & Scopus \\
V7  & Tax et al. (2017)              & PPM seminal (BPM baseline)         & Control \\
V8  & Rogge-Solti \& Weske (2015)    & Stochastic + PM forecasting        & Scopus \\
V9  & Rozinat et al. (2009)          & PM $\rightarrow$ Simulation bridge & Control \\
V10 & Gupta \& Sureka (2017)         & PM multi-repository in SE          & ACM \\
\bottomrule
\end{tabular}
\end{table}


\subsection{Inclusion and Exclusion Criteria}
\label{sec:slr-criteria}

The inclusion criteria are organized in two tiers. Formal criteria
(IC1--IC3) filter on publication attributes and are applied during the
database search configuration. Content criteria (IC4a--IC4d) require that the
study's research contribution addresses at least one of four specific
technique-domain intersections; they are applied during title/abstract and
full-text screening. Exclusion criteria (EC1--EC5) are applied cumulatively
across both screening phases. Criteria EC1--EC4 are evaluated by the
LLM-assisted screener; EC5 (full text inaccessible) is applied
programmatically after the inter-library-loan recovery pass for Band~D/E
papers fails to retrieve a copy.

\begin{table}[htbp]
\centering
\caption{Inclusion and exclusion criteria.}
\label{tab:criteria}
\small
\setlength{\tabcolsep}{4pt}
\begin{tabular}{c c p{10.2cm}}
\toprule
\textbf{Type} & \textbf{ID} & \textbf{Criterion} \\
\midrule
\multirow{7}{*}{Inclusion}
& IC1  & Publication period: 1994--2026 (from Whittaker \& Thomason onward). \\
& IC2  & Language: English. \\
& IC3  & Peer-reviewed journal article, conference paper, or workshop paper
          ($\geq$4 pages). \\
& IC4a & Applies process mining techniques (discovery, conformance checking,
          predictive process monitoring) to software development artifacts
          (commits, issues, pull requests, CI/CD logs, bug trackers, VCS). \\
& IC4b & Uses stochastic models (Markov chains, Monte Carlo, stochastic Petri
          nets, absorbing chains) to analyze software development processes or
          workflows. \\
& IC4c & Proposes or evaluates forecasting of software process metrics (lead
          time, cycle time, remaining time, throughput, defect rates) from
          event logs or repository data. \\
& IC4d & Mines software repositories (GitHub, Jira, VCS, CI/CD platforms) to
          discover, analyze, or improve software process models. \\
\midrule
\multirow{5}{*}{Exclusion}
& EC1  & Application domain is exclusively outside software development
          (healthcare, manufacturing, supply chain) without demonstrated
          relevance to SDLC processes. \\
& EC2  & Software appears only as the implementation platform; the process
          being analyzed is not a software development process. \\
& EC3  & Purely algorithmic or theoretical contribution without empirical
          evaluation in a software development context. \\
& EC4  & Secondary study (SLR, mapping study, survey) not specifically
          addressing PM or stochastic methods in software engineering. \\
& EC5  & Full text inaccessible despite all retrieval attempts (open access,
          institutional, ILL, author-provided preprint). Applied programmatically
          after Band~D/E recovery. \\
\bottomrule
\end{tabular}
\end{table}

At least one of IC4a--IC4d must be satisfied for inclusion. Criterion IC4b
ensures that seminal works applying Markov chains to software processes---such
as Whittaker and Thomason~\cite{whittaker1994}---are captured despite
predating the term ``process mining.''


\subsection{LLM-Assisted Screening Methodology}
\label{sec:slr-llm}

Given the scale of the working set (2{,}340 papers) and the requirement for
systematic, reproducible screening decisions with traceable rationale, the
title/abstract screening phase employs a language model as the primary
screener. Specifically, \textit{claude-haiku-4-5-20251001} is used via the
Anthropic Message Batches API, with a structured prompt encoding the IC4a--IC4d
and EC1--EC4 criteria (EC5 is applied programmatically post-screening), the PATHCAST scope description, and decision rules for
the three-class output: \texttt{include}, \texttt{exclude}, or \texttt{maybe}.

The screening prompt instructs the model to (i)~identify which ICs are
satisfied, (ii)~identify which ECs apply, (iii)~assign a decision with a
one-to-two sentence rationale, and (iv)~return the result as structured
JSON. The \texttt{maybe} class captures genuine uncertainty---typically
arising from missing abstracts or ambiguous domain terminology---without
forcing a binary decision.

At the full-text screening phase, the same model is applied with a stricter
prompt that enforces a binary outcome (\texttt{include} or \texttt{exclude})
with \texttt{pending} reserved only when the abstract is genuinely
insufficient for a decision.

This approach extends prior work on LLM-assisted systematic reviews
\cite{syriani2023,khraisha2024} to the software process mining domain. The
primary screener produces structured outputs that enable post-hoc quality
analysis. Validation follows a two-stage protocol: (i)~all \texttt{include}
decisions are manually verified by the author; (ii)~a stratified random 20\%
sample of T/A and FT decisions is independently re-rated by a different model
family (\textit{claude-sonnet-4-6}) acting as a second rater, and inter-rater
agreement is reported via Cohen's $\kappa$ following the Landis and Koch
(1977) interpretation thresholds. The substantial-agreement threshold
($\kappa \geq 0.61$) is the target reported by Khraisha~et~al.~\cite{khraisha2024}
for LLM-assisted screening in the medical SLR domain and adopted here for
SE-domain SLRs. Cross-model verification is recommended by
Syriani~et~al.~\cite{syriani2023} as a practical proxy when a fully
independent human second rater is not feasible at the working-set scale
considered here. Results of this verification are reported in
Section~\ref{sec:slr-threats} (Internal Validity).


\subsection{Quality Assessment}
\label{sec:slr-qa}

Each included study is assessed against a quality checklist adapted from
Dyb{\aa} and Dings{\o}yr~\cite{dyba2008}, with eight binary criteria:

\begin{table}[htbp]
\centering
\caption{Quality assessment checklist.}
\label{tab:qa}
\begin{tabular}{clc}
\toprule
\textbf{\#} & \textbf{Criterion} & \textbf{Score} \\
\midrule
QA1 & Are the research objectives clearly stated?                    & 0/1 \\
QA2 & Is the software engineering context described in detail?       & 0/1 \\
QA3 & Is the data source (event log) described reproducibly?         & 0/1 \\
QA4 & Is the PM or stochastic technique formally defined?            & 0/1 \\
QA5 & Are results validated empirically?                             & 0/1 \\
QA6 & Are threats to validity discussed?                             & 0/1 \\
QA7 & Is the study reproducible (data and/or code available)?        & 0/1 \\
QA8 & Are process model quality metrics reported (fitness, precision)?& 0/1 \\
\bottomrule
\end{tabular}
\end{table}

Studies scoring below 4.0/8 (50\%) are excluded from the synthesis. QA8
reflects the centrality of process model quality in the PATHCAST pipeline.

QA scoring is performed by the same LLM-assisted protocol used for screening
(Section~\ref{sec:slr-llm}): \textit{claude-haiku-4-5-20251001} receives a
strict rubric for each criterion and returns a binary score with a brief
justification. The full prompt and the per-study scores are part of the
replication package (Section~\ref{sec:slr-replication}).
Table~\ref{tab:qa-results} reports the quality-assessment outcome for the
complete confirmed set ($n = 169$). All 169 studies were scored.

\input{results/qa_assessment_summary}

The 33 studies below the QA threshold are predominantly short workshop papers,
position papers, or extended abstracts that fail QA3 (reproducible data
source) and QA4 (formally defined technique) simultaneously. They are not
removed from the PRISMA flow; they are retained in the included set as
context but are not used in the F1--F5 evidence base reported in
Section~\ref{sec:slr-gaps}.


\subsection{Data Extraction}
\label{sec:slr-extraction}

For each included study, the attributes in Table~\ref{tab:extraction} are
extracted.

\begin{table}[htbp]
\centering
\caption{Data extraction form.}
\label{tab:extraction}
\small
\setlength{\tabcolsep}{4pt}
\begin{tabular}{p{3.6cm} p{8.1cm}}
\toprule
\textbf{Field} & \textbf{Description} \\
\midrule
ID                           & Unique identifier \\
Authors, Year, Venue         & Bibliographic metadata \\
SDLC Phase(s)                & Lifecycle phases covered \\
Event Log Source             & Jira, GitHub, GitLab, Jenkins, etc. \\
Event Log Construction       & Manual, semi-automatic, or automatic \\
PM Technique Category        & Discovery / Conformance / Enhancement / Prediction \\
Specific Algorithm(s)        & Alpha, Heuristics, Inductive Miner, etc. \\
Stochastic Method            & Markov / Monte Carlo / Stochastic Petri Net / None \\
ML Technique                 & RF, XGBoost, LSTM, etc. / None \\
Prediction Target            & Remaining time, outcome, lead time, delay, etc. \\
Integration Level            & L0 None / L1 Sequential / L2 Pipeline / L3 Unified \\
Validation Type              & Case study / Experiment / Industrial / Simulation \\
Tool/Platform                & PM4Py, ProM, Disco, custom, etc. \\
Dataset Size                 & Number of events and cases analyzed \\
Process Model Quality        & Fitness, precision, generalization (if reported) \\
\bottomrule
\end{tabular}
\end{table}


\section{Search Results and Study Selection}
\label{sec:slr-results-selection}

\subsection{Database Retrieval and Deduplication}

The three search strings executed across five databases returned 7{,}740
records. Backward and forward snowballing from the control set yielded an
additional 595 records. Together with 12 control papers maintained as a
validation-only set, the initial corpus totaled 8{,}347 records.
Table~\ref{tab:retrieval} summarizes the retrieval by source.

\begin{table}[htbp]
\centering
\caption{Record retrieval by source.}
\label{tab:retrieval}
\begin{tabular}{lrr}
\toprule
\textbf{Source} & \textbf{Raw Records} & \textbf{After Internal Dedup} \\
\midrule
Scopus (6 queries)      & 2{,}377 & 1{,}852 \\
SpringerLink (10 exports) & 2{,}884 & 1{,}543 \\
IEEE Xplore (manual)    & 1{,}386 & 1{,}263 \\
Snowballing             &   595   &   575   \\
Web of Science (4 queries) & 997  &   514   \\
ACM Digital Library (3 queries) & 96 &  34  \\
Control set (validation only) & 12 &    2   \\
\midrule
\textbf{Total} & \textbf{8{,}347} & \textbf{5{,}783} \\
\bottomrule
\end{tabular}
\end{table}

Cross-source deduplication removed 2{,}564 duplicate records (30.7\%),
yielding 5{,}781 unique primary-corpus papers plus 2 unique control papers.
Deduplication used normalized DOI matching, followed by fuzzy title matching
(Levenshtein similarity $\geq 0.9$) for records without DOIs.

\subsection{Operational Working Set Construction}

Screening 5{,}781 records in a single pass would require substantial
resources. A working set of 2{,}340 papers was therefore constructed from the
highest-signal primary queries---those whose query design directly targets the
PATHCAST intersection (process mining, stochastic modeling, and software
development)---as shown in Table~\ref{tab:working-set}. The remaining 3{,}441
records constitute a \textit{high-recall auxiliary corpus}, retained as a
frozen baseline and available for supplementary analysis if gaps in the
primary results warrant expansion.

\begin{table}[htbp]
\centering
\caption{Working set composition.}
\label{tab:working-set}
\small
\setlength{\tabcolsep}{4pt}
\begin{tabular}{p{2.2cm} p{8.3cm} r}
\toprule
\textbf{Database} & \textbf{Queries included} & \textbf{Unique papers} \\
\midrule
Scopus & scopus\_principal, scopus\_complementar, scopus\_fronteira,
         scopus\_markov\_testing, scopus\_stochastic\_petri,
         scopus\_simulation\_bridge & 2{,}197 \\
IEEE Xplore & ieee\_manual\_01 (PM + SE subgroup) & 97 \\
ACM DL & acm\_principal, acm\_msr & 46 \\
\midrule
\textbf{Total} & & \textbf{2{,}340} \\
\bottomrule
\end{tabular}
\end{table}

The working set preserves $10/10$ control papers when the validation-only
control set is included, and $8/10$ from primary queries alone---both above
the 80\% threshold.

\subsection{Title/Abstract Screening}

All 2{,}340 working-set papers were screened against IC4a--IC4d and EC1--EC4
using the LLM-assisted protocol described in Section~\ref{sec:slr-llm}. The
screening was executed across five sequential batches of up to 500 papers each
via the Anthropic Batches API, producing structured JSON outputs with decision,
rationale, and matched criteria for each record.

\begin{table}[htbp]
\centering
\caption{Title/abstract screening results.}
\label{tab:ta-screening}
\begin{tabular}{lrr}
\toprule
\textbf{Decision} & \textbf{Count} & \textbf{\%} \\
\midrule
Include & 111 & 4.7 \\
Maybe   & 775 & 33.1 \\
Exclude & 1{,}454 & 62.2 \\
\midrule
\textbf{Total} & \textbf{2{,}340} & 100.0 \\
\bottomrule
\end{tabular}
\end{table}

The \texttt{include} set (111 papers) and the \texttt{maybe} set (775 papers)
together form the full-text review queue of 886 papers. The
\texttt{exclude} decisions were driven primarily by EC1 (domain outside
software development, principally generic business process management) and EC3
(theoretical algorithm proposals without SE evaluation).

\subsection{Full-Text Screening}

The 886 papers forwarded from the T/A phase were subjected to LLM-assisted
full-text screening using the strict prompt described in
Section~\ref{sec:slr-llm} (decisions limited to \texttt{include},
\texttt{exclude}, or \texttt{pending}; \texttt{maybe} is not permitted at this
stage). To maximise the number of papers assessable by the LLM, a
multi-source abstract-enrichment cascade was executed prior to
screening---combining DOI-based lookups via Semantic Scholar, OpenAlex, CORE,
and Crossref, followed by title-based fallback lookups---raising abstract
coverage from 16.6\% to 72.6\% (643 of 886 papers).\footnote{Papers were
  processed in order of a five-band priority scheme: Band~A (91 papers,
  T/A \texttt{include} $\geq$2 ICs), Band~B (20, T/A \texttt{include} 1~IC),
  Band~C (39, T/A \texttt{maybe} + abstract), Band~D (152, T/A \texttt{maybe}
  + IC signal, no abstract), Band~E (584, T/A \texttt{maybe}, no IC signal, no
  abstract). Abstract enrichment and LLM re-screening were applied to Bands~A,
  B, C, and the subset of D/E that gained an abstract after enrichment.}

Table~\ref{tab:ft-screening} summarises the outcome.

\begin{table}[htbp]
\centering
\caption{Full-text screening outcomes (886 papers forwarded from T/A phase).}
\label{tab:ft-screening}
\small
\setlength{\tabcolsep}{4pt}
\begin{tabular}{lrr}
\toprule
\textbf{Outcome} & \textbf{Count} & \textbf{\%} \\
\midrule
Included                             & 169 & 19.1 \\
Excluded                             & 717 & 80.9 \\
\quad EC3 (algorithmic, no SE eval.) & (229) & (25.8) \\
\quad EC1 (non-SE domain)            & (205) & (23.1) \\
\quad EC5 (full text inaccessible)   & (148) & (16.7) \\
\quad EC2 (SW as implementation)     & (110) & (12.4) \\
\quad EC4 (out-of-scope secondary)   & (60)  & (6.8)  \\
\midrule
\textbf{Total} & \textbf{886} & 100.0 \\
\bottomrule
\end{tabular}

\smallskip
\footnotesize\textit{Note:} EC counts are non-exclusive --- a single paper may
trigger multiple ECs when more than one applies (e.g., a non-SE-domain paper
whose contribution is purely theoretical). The sum of per-EC counts therefore
exceeds the 717 excluded papers; the percentages above are with respect to the
886-paper FT queue.
\end{table}

Among the 717 excluded papers, EC3 (algorithmic or theoretical contribution
without SE evaluation, $n=229$) and EC1 (domain outside software development,
$n=205$) jointly account for the largest share of FT-stage exclusions. EC5
(full text inaccessible despite all retrieval attempts, $n=148$) corresponds
to the Band~D/E subset for which no open-access copy, institutional access,
or author-provided preprint could be located; these are recorded in a
dedicated traceability file and documented in the PRISMA flow as a named
exclusion stratum. EC2 (software as implementation platform rather than
analyzed process, $n=110$) and EC4 (out-of-scope secondary studies, $n=60$)
account for the remainder.

Figure~\ref{fig:prisma} presents the complete PRISMA 2020 flow
diagram~\cite{page2021} of the selection process.

\begin{figure}[p]
\centering
\begin{tikzpicture}[
  %% Main flow boxes (centre column) --- progressively narrower for funnel feel
  stg/.style={draw, rounded corners=4pt, align=center,
              text width=#1, minimum height=1.1cm,
              font=\small, fill=white, thick},
  stg/.default=8cm,
  %% Exclusion / note boxes (right column)
  exc/.style={draw, rounded corners=4pt, align=left,
              inner xsep=5pt, inner ysep=4pt,
              text width=5.0cm, minimum height=0.9cm,
              font=\footnotesize, fill=gray!8, draw=gray!60},
  %% Phase labels (left column)
  lbl/.style={font=\footnotesize\bfseries, text=gray!70,
              rotate=90, anchor=center},
  arr/.style={->, >=Stealth, thick},
  sidearr/.style={->, >=Stealth, dashed, draw=gray!70, thick},
]

%% -- PHASE: IDENTIFICATION ----------------------------------------------

\node[lbl] at (-6.8,-1.1) {IDENTIFICATION};
\draw[gray!40, dashed] (-6.0,-2.4) -- (-6.0, 0.2);

\node[stg=9.5cm, minimum height=1.3cm] (id) at (-0.5,0)
  {\textbf{Records identified} ($n = 8{,}347$)\\[3pt]
   Scopus~2{,}377 \textbullet{} \mbox{SpringerLink}~2{,}884\\
   IEEE~1{,}386 \textbullet{} WoS~997 \textbullet{} ACM~96\\[2pt]
   Snowballing~595 \textbullet{} Control~12};

\node[stg=9.5cm] (dedup) at (-0.5,-2.2)
  {\textbf{After cross-source deduplication} ($n = 5{,}783$)\\[3pt]
   2{,}564 duplicates removed (30.7\%)};

\draw[arr] (id) -- (dedup);

%% -- PHASE: SCREENING ----------------------------------------------------

\node[lbl] at (-6.8,-5.6) {SCREENING};
\draw[gray!40, dashed] (-6.0,-2.4) -- (-6.0,-8.0);

\node[stg=7.5cm, minimum height=1.2cm] (ws) at (-0.5,-4.5)
  {\textbf{Operational working set} ($n = 2{,}340$)\\[3pt]
   Scopus~2{,}197 \textbullet{} IEEE~97 \textbullet{} ACM~46};

\node[exc] (wsdrop) at (6.0,-4.5)
  {Deferred to auxiliary\\high-recall corpus ($n = 3{,}441$):\\
   \mbox{SpringerLink}, WoS,\\Snowball, extra IEEE/ACM};

\draw[arr] (dedup) -- (ws);
\draw[sidearr] (ws.east) -- (wsdrop.west);

\node[stg=7.5cm, minimum height=1.2cm] (ta) at (-0.5,-7.0)
  {\textbf{Title/Abstract screened} ($n = 2{,}340$)\\[3pt]
   LLM-assisted (Claude Haiku 4.5)};

\node[exc] (taexc) at (6.0,-7.0)
  {Excluded at T/A ($n = 1{,}454$)\\
   EC1: non-SE domain (BPM)\\
   EC3: theoretical, no SE eval.};

\draw[arr] (ws) -- (ta);
\draw[sidearr] (ta.east) -- (taexc.west);

%% -- PHASE: ELIGIBILITY --------------------------------------------------

\node[lbl] at (-6.8,-10.8) {ELIGIBILITY};
\draw[gray!40, dashed] (-6.0,-8.0) -- (-6.0,-13.6);

\node[stg=7cm, minimum height=1.3cm] (ftq) at (-0.5,-9.4)
  {\textbf{Full-text review queue} ($n = 886$)\\[3pt]
   T/A include~111 \textbullet{} T/A maybe~775\\
   Bands~A--E by priority};

\draw[arr] (ta) -- (ftq);

\node[stg=7cm, minimum height=1.4cm] (ft) at (-0.5,-12.2)
  {\textbf{Full-text screened} ($n = 886$)\\[3pt]
   LLM + manual \textbullet{} abstract coverage 72.6\% after enrichment\\
   Screen-blanks pass (186 FT-undecided papers)};

\node[exc] (ftexc) at (6.0,-11.6)
  {Excluded at FT ($n = 717$)\\
   EC3: 229 theoretical, no SE\\
   EC1: 205 non-SE domain\\
   EC5: 148 FT inaccessible\\
   EC2: 110 \textbullet{} EC4: 60};

\draw[arr] (ftq) -- (ft);
\draw[sidearr] (ft.east) -- (ftexc.west);

%% -- PHASE: INCLUDED ----------------------------------------------------

\node[lbl] at (-6.8,-14.6) {INCLUDED};
\draw[gray!40, dashed] (-6.0,-13.6) -- (-6.0,-15.6);

\node[stg=8cm, minimum height=1.3cm] (inc) at (-0.5,-14.8)
  {\textbf{Studies included} ($n = 169$)\\[3pt]
   75 with PDF \textbullet{} LLM extraction complete (169/169)};

\draw[arr] (ft) -- (inc);

\end{tikzpicture}
\caption{PRISMA 2020 flow diagram of the study selection process.
  Phase labels follow the four-stage structure: Identification, Screening,
  Eligibility, and Included.}
\label{fig:prisma}
\end{figure}


\section{Results and Synthesis}
\label{sec:slr-synthesis}

\subsection{Overview of Included Studies}

The confirmed set of 169 primary studies spans three decades
(1995--2026), with a clear acceleration from 2014 onward
(Figure~\ref{fig:year-dist}). Seventy-one percent of confirmed papers were
published between 2015 and 2026, reflecting the maturation of open-source
process mining toolkits (PM4Py, ProM) and the broader adoption of deep
learning frameworks in software engineering research.

\begin{figure}[htbp]
\centering
\begin{tikzpicture}
\begin{axis}[
  ybar,
  bar width=0.34cm,
  width=14cm, height=5cm,
  xlabel={Year},
  ylabel={Papers},
  xtick=data,
  xticklabels={1995,1998,2000,2004,2005,2006,2007,2008,2009,2010,2011,2012,2013,2014,2015,2016,2017,2018,2019,2020,2021,2022,2023,2024,2025,2026},
  x tick label style={rotate=45, anchor=east, font=\scriptsize},
  ymin=0, ymax=17,
  ytick={0,2,4,6,8,10,12,14,16},
  enlarge x limits=0.03,
  fill=gray!40,
  draw=black,
  nodes near coords,
  nodes near coords style={font=\tiny, above},
]
\addplot coordinates {
  (1,1)(2,1)(3,1)(4,2)(5,1)(6,5)(7,3)(8,3)(9,5)(10,3)(11,10)(12,5)(13,1)(14,11)(15,12)(16,9)(17,13)(18,10)(19,12)(20,7)(21,15)(22,11)(23,10)(24,7)(25,8)(26,3)
};
\end{axis}
\end{tikzpicture}
\caption{Distribution of confirmed studies by publication year ($n = 169$).}
\label{fig:year-dist}
\end{figure}

By document type, the majority are conference papers and journal articles.
By source database, Scopus is the dominant contributor, followed by
IEEE~Xplore and ACM~DL. Data extraction was completed for all 169 studies
using a structured LLM-assisted pipeline (\textit{claude-haiku-4-5-20251001})
operating on locally available PDFs (75/169, of which 73 were operationally
coded in the triangulation subset) or abstract/title metadata (94/169).

\subsection{Inclusion Criterion Coverage}

Figure~\ref{fig:ic-dist} shows the distribution of inclusion criteria among
the 169 confirmed papers (criteria are non-exclusive; a paper may satisfy
multiple).

\begin{figure}[htbp]
\centering
\begin{tikzpicture}
\begin{axis}[
  xbar,
  bar width=0.55cm,
  width=10cm, height=4.5cm,
  xlabel={Papers matching criterion},
  symbolic y coords={IC4 (MSR for process), IC3 (Forecasting), IC2 (Stochastic), IC1 (PM in SE)},
  ytick=data,
  y tick label style={font=\small},
  xmin=0, xmax=115,
  nodes near coords,
  nodes near coords align={horizontal},
  fill=gray!40,
  draw=black,
  enlarge y limits=0.25,
]
\addplot coordinates {
  (121,{IC1 (PM in SE)})
  (52,{IC2 (Stochastic)})
  (29,{IC3 (Forecasting)})
  (45,{IC4 (MSR for process)})
};
\end{axis}
\end{tikzpicture}
\caption{Inclusion criteria matched by the confirmed set ($n = 169$;
  non-exclusive; IC labels map to IC4a--IC4d in the protocol).}
\label{fig:ic-dist}
\end{figure}

IC1/IC4a (process mining applied to SE artifacts) is the dominant criterion,
matched by 121 papers (71.6\%). IC4/IC4d (repository mining for process modeling)
is second at 45 papers (26.6\%), reflecting the MSR community's overlapping
contributions. IC2/IC4b (stochastic modeling) appears in 52 papers (30.8\%) and
IC3/IC4c (forecasting of process metrics) in 29 papers (17.2\%). Notably, 68
papers (40.2\%) match IC1 only, without stochastic or forecasting components,
providing direct empirical evidence for the ``Descriptive Ceiling'' identified
in Finding~F1 below.

\subsection{Prevalent Exclusion Patterns}

Among the 717 FT-excluded papers (Table~\ref{tab:ft-screening}), the largest
exclusion reason is EC3 (algorithmic or theoretical contribution without
software-engineering evaluation, $n=229$), indicating a substantial body of
process mining and stochastic modeling research that has not been empirically
applied to software engineering contexts. EC1 (domain exclusively outside
software development, $n=205$) is the second-largest reason and consists
primarily of generic business process management (BPM) studies that use the
same terminology as the search strings but apply techniques to organizational
processes unrelated to SDLC. EC5 (full text inaccessible, $n=148$) closes the
group of large strata, followed by EC2 (software as implementation platform
rather than analyzed process, $n=110$) and EC4 (out-of-scope secondary study,
$n=60$). EC counts are non-exclusive: a single paper may trigger more than
one EC. The dominance of EC3 and EC1 directly supports Findings F3 and F4
below: process mining and stochastic methods are widely studied in the
abstract or in non-SE domains, but their application to SDLC processes
remains comparatively narrow.

\subsection{Full-Text Extraction and PDF-Based Triangulation}
\label{sec:slr-pdf-enrichment}

Structured data extraction was performed on the full 169-study set using the
extraction template defined in Table~\ref{tab:extraction}. For each study,
the LLM extracted the research question, study type, research contribution,
PM technique, stochastic technique, software artifact, software process, data
source, dataset availability, tool used, main finding, limitations, and
replication-package status. Title-and-abstract metadata was used as the
primary input for studies without locally available PDFs (94/169); for the
remaining 75/169 studies a locally accessible PDF was available and used to
triangulate against the LLM-extracted fields. Of these 75 PDFs, 73 were
fully coded in the operational triangulation spreadsheet used for the
figures and positioning tables below.

The 73-paper PDF triangulation subset is therefore not a separate selection
stage and does not override the PRISMA counts. It is used in this chapter as
a verification layer for the LLM-extracted fields used in Findings F1--F5.

Figure~\ref{fig:pdf-decision-dist} summarizes the operational triage profile
of the 73 coded PDFs (54 confirm-include, 13 confirm-include with
caveats, 6 confirm-exclude after full-text inspection); the formal selection
decisions reported in the PRISMA flow (Section~\ref{sec:slr-prisma}) are
unchanged.

\begin{figure}[tbp]
\centering
\begin{tikzpicture}
\begin{axis}[
  ybar,
  bar width=0.7cm,
  width=9cm, height=5cm,
  ylabel={Papers},
  symbolic x coords={Include,Maybe,Exclude},
  xtick=data,
  ymin=0, ymax=60,
  ytick={0,10,20,30,40,50,60},
  nodes near coords,
  nodes near coords style={font=\small, above},
  fill=gray!45,
  draw=black,
  enlarge x limits=0.25,
]
\addplot coordinates {(Include,54) (Maybe,13) (Exclude,6)};
\end{axis}
\end{tikzpicture}
\caption{Operational decision distribution in the 73-paper PDF triangulation subset.}
\label{fig:pdf-decision-dist}
\end{figure}

\subsection{RQ1: Application Landscape}
\label{sec:slr-rq1}

\subsubsection{RQ1.1: SDLC Phases Covered}

The structured extraction over the full 169-study set confirms a strongly
uneven coverage of the SDLC. \textbf{Development activities} (coding,
commits, pull requests, repository evolution) dominate at 133 papers
(78.7\%); \textbf{quality-related activities} (testing, defect handling, code
review, process evaluation) appear in 74 papers (43.8\%);
\textbf{maintenance} appears in 38 papers (22.5\%);
\textbf{requirements engineering} in only 22 papers (13.0\%); and
\textbf{delivery/DevOps contexts} (CI/CD, build and release pipelines) in
just 18 papers (10.7\%). \textbf{Design and architecture} are not the primary
focus of any extracted study, and only 3 papers (1.8\%) center the analysis
on project management.

Two conclusions follow. First, the literature is concentrated on phases where
digital traces are abundant and already timestamped, such as commits, issues,
test events, and pipeline executions. Second, the early SDLC phases remain
structurally underrepresented, which weakens the ability of current
approaches to support end-to-end forecasting from requirements to delivery.
For PATHCAST, this supports a pragmatic focus on operational phases with
reliable event data while leaving room for future extensions toward
requirements and design logs.

\subsubsection{RQ1.2: Event Log Sources and Construction}

Across the 169-study set, event logs are constructed primarily from four
source families: \textbf{version-control repositories} (Git/GitHub/SVN
commits) appear in 84 papers (49.7\%); \textbf{issue tracking systems}
(especially Jira and GitHub Issues) in 37 papers (21.9\%); \textbf{IDE
interaction logs} in 22 papers (13.0\%); and \textbf{CI/CD or build
pipelines} in 15 papers (8.9\%). The common pattern is not a native event
log but rather a \emph{derived} event log assembled from repository metadata,
commit histories, issue transitions, and (where available) pipeline records.

The dominant construction problem is case correlation across tools. Studies
frequently analyze one source in isolation or perform ad hoc joins between
issues and commits, which limits reproducibility. This confirms that event-log
construction is not a preparatory detail but a central methodological layer in
software process mining. The finding directly justifies Stage~1 of PATHCAST as
an explicit transformation pipeline rather than an implicit data-cleaning step.

A non-trivial fraction of studies report data sources only in coarse terms
(\textit{repository data}, \textit{VCS history}, \textit{open-source
projects}) without naming specific projects, time windows, or extraction
queries. Across the 169-study set, only 25 papers (14.8\%) explicitly state
that their dataset is publicly available and only 4 papers (2.4\%) declare a
replication package. This itself is evidence of incomplete reporting in the
literature and provides direct empirical grounding for Finding~F2 below.

A complementary body of work constructs Markov models directly from GitHub
repository traces, bypassing explicit process discovery in favour of manual
state-space definition. Jo et al.\ \cite{jo2023} modelled developer activities
in GitHub repositories as Discrete-Time Markov Chains (DTMCs), finding that
commits account for 62.73\% of long-term activity, and employed probabilistic
temporal logic to extract collaboration patterns relevant to project management.
This approach was extended in \cite{jo2024}, where Probabilistic Model Checking
with Computation Tree Logic was applied across five repositories to characterise
temporal dynamics, state-transition probabilities, and key behavioural drivers,
yielding transparent and explainable project forecasts. Ortu et al.\ \cite{ortu2023}
complemented these structural analyses by combining Markov chains with social
network analysis to identify fault-insertion and fault-fixing behavioural patterns
across Apache Software Foundation projects, demonstrating that developer expertise
significantly moderates defect introduction likelihood. Collectively, these studies
confirm that SDLC event logs extracted from public repositories are sufficiently
rich to support state-transition modelling, satisfying a foundational prerequisite
for PATHCAST Stage~1. However, the event-to-state mapping adopted across these
works remains unstandardised and ad hoc, representing a methodological gap that
PATHCAST's process mining discovery stage aims to address systematically.


\subsection{RQ2: Techniques and Prediction}
\label{sec:slr-rq2}

\subsubsection{RQ2.1: Discovery Algorithms Applied to SDLC}

The canonical origin of process mining applied to software development contexts
can be traced directly to the foundational contributions of Cook and Wolf in the
mid-1990s. In their seminal work, \cite{cook1995} introduced three automated
process discovery methods---algorithmic grammar inference, Markov models, and
neural networks---demonstrating that finite-state machine representations of
software processes could be extracted automatically from event traces,
substantially lowering the barrier to formal process modeling. This line of
inquiry was consolidated in \cite{cook1998}, which refined the event-based
discovery methodology and established the conceptual vocabulary of inferring
workflow structures from historical execution data, a paradigm that remains
central to the field today. Crucially, both contributions embedded stochastic
reasoning---particularly Markov chain representations---within the discovery
pipeline, anticipating the probabilistic extensions that would later characterise
mature process mining frameworks. The translation of these principles into an
explicit, named framework for software engineering contexts was accomplished by
\cite{rubin2007}, who applied the ProM toolkit to source code management logs
and demonstrated multi-perspective analysis spanning control-flow, informational,
and organisational dimensions. Together, these three works constitute the seminal
arc of IC1, establishing automated discovery from SDLC event logs as both
technically feasible and analytically productive, and providing PATHCAST with
its foundational methodological lineage.

The complete extracted set confirms that \textbf{process mining remains the
dominant analytical family}: 121 of 169 studies (71.6\%) match IC1 in the
form of discovery, conformance checking, or process-oriented repository
analysis. At the technique level, the most frequently named PM algorithms
across the 169 studies are the \textbf{inductive miner} (51 papers),
\textbf{alpha miner} (45 papers), \textbf{conformance checking} (30 papers),
and \textbf{heuristic miner} (22 papers); 11 papers use social network
analysis on developer interactions. The literature is markedly richer in
\emph{process representation} than in probabilistic forecasting: only 29
papers (17.2\%) match IC3 (forecasting of process metrics), and the dominant
contribution type is discovery or conformance checking rather than predictive
analytics. This is the transition that PATHCAST targets: moving from
discovered process structures to state-based probabilistic forecasting.

The stochastic-technique distribution at the same 169-study level shows the
same fragmentation. Only 52 of 169 studies (30.8\%) name an explicit
stochastic technique: 38 papers (22.5\%) use Markov chains, 15 (8.9\%) use
Monte Carlo simulation, 9 (5.3\%) use stochastic Petri nets, and 8 (4.7\%)
use system dynamics. The remaining studies either do not adopt a stochastic
component (110 papers) or place stochastic methods in an auxiliary role
(\textit{other}, 29 papers). The forecasting literature thus remains
fragmented across adjacent technique families rather than consolidated around
a single process-aware predictive paradigm, as shown in
Figure~\ref{fig:pdf-technique-dist}.

\begin{figure}[tbp]
\centering
\begin{tikzpicture}
\begin{axis}[
  xbar,
  bar width=0.45cm,
  width=12cm, height=5.5cm,
  xlabel={Papers},
  symbolic y coords={Process Mining,Markov Chain,Simulation,Stochastic Petri Net,System Dynamics},
  ytick=data,
  xmin=0, xmax=130,
  nodes near coords,
  nodes near coords align={horizontal},
  fill=gray!45,
  draw=black,
  enlarge y limits=0.18,
]
\addplot coordinates {
  (121,{Process Mining})
  (38,{Markov Chain})
  (15,{Simulation})
  (9,{Stochastic Petri Net})
  (8,{System Dynamics})
};
\end{axis}
\end{tikzpicture}
\caption{Main technique families identified in the 169-study confirmed set.
Counts are non-exclusive: a paper may use multiple technique families.}
\label{fig:pdf-technique-dist}
\end{figure}

\subsubsection{RQ2.2: Predictive and Stochastic Techniques}

Three distinct sub-families of stochastic techniques emerge from the confirmed
literature. The first, \textbf{Markov-chain-based prediction}, encompasses
models that represent software processes as discrete state transitions, including
reinforcement learning agents grounded in Markov Decision Processes for
pull-request outcome prediction \cite{joshi2024}, probabilistic risk
quantification from repository data \cite{jeon2015}, defect removal
effectiveness under code churn \cite{tian2005}, agile state-space planning and
tracking \cite{tyagi2021}, release-time forecasting via generalised reliability
models \cite{washizaki2015}, and Stochastic Automata Networks for Waterfall team
estimation \cite{massitela2018}. The second sub-family, \textbf{stochastic
Petri nets}, models concurrency and asynchrony in DevOps and CI/CD pipelines
through non-Markovian specifications \cite{benmesmia2021} and sensitivity
analyses that identify bottleneck stages governing delivery probability
\cite{bhadra2022,bhadra2023}. The third sub-family, \textbf{Monte Carlo
simulation}, drives project-level forecasting in hybrid System Dynamics models
\cite{li2020}, agile risk assessment frameworks \cite{lunesu2021}, cycle-time
economic analysis \cite{magennis2015}, and the PRIMAD digital-twin architecture
for SDLC \cite{guinea2025}.

Despite their methodological diversity, all three sub-families share a common
parametrisation strategy that constitutes a structural limitation. Transition
probabilities, firing rates, and distributional parameters are invariably derived
from aggregate project statistics---historical defect counts, phase-level duration
averages, or summary velocity metrics---rather than from transition frequencies
extracted through systematic process mining of fine-grained event logs
\cite{nafreen2020,bhadra2022,li2020}. Consequently, model states remain anchored
to coarse lifecycle phases and cannot reflect the richer, empirically grounded
activity sequences that event logs encode. This observation constitutes
\textbf{Finding~F3}: \emph{stochastic forecasting models for software development
are parametrised from aggregate statistics rather than from mined process
structures, confining their state spaces to coarse lifecycle granularity and
decoupling model topology from observed execution behaviour}.

Three studies exemplify the productive convergence of process mining and
stochastic modelling at the component level. Incerto~\cite{incerto2025}
demonstrated that variable-length Markov chains outperform eighteen existing
stochastic conformance-checking techniques across ten of eleven benchmark
datasets, establishing that higher-order sequential dependencies in event logs
can be captured without sacrificing computational tractability.
L\'opez-Pintado~\cite{lopezpintado2023} extended process-simulation fidelity by
replacing deterministic resource-availability calendars with probabilistic
counterparts discovered directly from event logs, showing that stochastic
resource representations more accurately replicate empirical cycle-time
distributions. Jalote~\cite{jalote2021} took a complementary perspective,
modelling programmer task sequences as first-order Markov chains derived from
IDE interaction logs and demonstrating that transferring the process patterns of
high-productivity developers to average-productivity peers yields measurable
productivity gains. Together, these contributions confirm that stochastic
extensions of process models improve both explanatory and predictive validity;
however, each study confines its stochastic layer to a single analytical
concern without integrating multiple stages into a unified forecasting pipeline.

The closest published work to a fully integrated multi-stage architecture is
PRIMAD, proposed by Guinea-Cabrera~\cite{guinea2025}, which orchestrates
pm4py-based process discovery, Akka actor-system simulation, and LightGBM
supervised learning within an evolutionary Digital Twin framework for the SDLC.
Nevertheless, PRIMAD remains at what may be characterised as Level~2 integration:
validation is reported per component rather than across the assembled pipeline,
no formal inter-stage contracts govern the handoffs between discovery,
simulation, and learning modules, and the architecture incorporates no dedicated
Monte Carlo forecasting layer capable of propagating uncertainty estimates
end-to-end. The present review therefore concludes that the Level~3 integration
niche---a formally contracted, end-to-end pipeline combining process discovery,
absorbing Markov chain construction, Monte Carlo simulation, and residual
correction---remains unoccupied in the extant literature.

Studies that combine process mining with forecasting objectives (IC1~$\cap$~IC3)
tend to approach prediction through \emph{process enhancement} rather than
through explicit stochastic modelling. Gupta~\cite{gupta2017} demonstrated that
combining inductive miner discovery with Markov chain-based predictive analytics
enables identification of recurring failure patterns in software maintenance
workflows. Pourbafrani et al.~\cite{pourbafrani2021} presented SIMPT, a tool
that extracts process trees with time parameters from event logs and enables
interactive what-if simulation; its successor work~\cite{buliga2025} further
integrates generative AI with business process simulation to produce
conformance-preserving what-if scenarios. Caldeira et al.~\cite{caldeira2022}
mine IDE interaction logs and show that process complexity metrics derived from
ProM correlate moderately ($r \approx 0.43$) with software cyclomatic
complexity, achieving predictive accuracies exceeding 92\%, demonstrating the
feasibility of process-derived features for product-quality prediction.

Taken together, this cluster confirms that PM and prediction are being connected
in the SE literature, but always at the level of sequential application---PM
output feeds a downstream model---rather than through a unified pipeline with
formal contracts. Crucially, none of these systems can express, as a first-class
output, the probability that a given process instance will reach completion
within $k$ activity cycles---the precise forecasting primitive that distinguishes
absorbing Markov chain formulations from ad hoc simulation runs. This gap
motivates PATHCAST's design, wherein the mined process structure feeds directly
into an absorbing Markov chain whose transient-state absorption probabilities
constitute a formally defined, queryable forecasting interface.


\subsection{RQ3: Integration and Gaps}
\label{sec:slr-rq3}

\subsubsection{RQ3.1: PM and Stochastic Integration Level}

Mapping the confirmed papers onto the L0--L3 integration scale reveals a clear
stratification. The largest cluster, comprising papers that touch IC2 and IC3
without IC1---including stochastic Petri net models of CI/CD pipelines
\cite{bhadra2022,bhadra2023}, Markov-based project estimation
\cite{massitela2018,tyagi2021}, and Monte Carlo risk simulation
\cite{lunesu2021,magennis2015}---operates almost entirely at \textbf{L1},
applying stochastic modelling to software processes without any process
discovery layer. Papers spanning IC1 and IC2, such as the VLMC conformance
checker of \cite{incerto2025} and the probabilistic resource-calendar discoverer
of \cite{lopezpintado2023}, reach \textbf{L2} by coupling discovered process
structure with stochastic reasoning, yet neither closes the loop toward
distributional schedule forecasting. Papers combining IC1 with IC3, such as
\cite{gupta2017} and \cite{caldeira2022}, use process mining descriptively to
assess workflows but lack the stochastic machinery required for Monte Carlo
distributional forecasts. The PRIMAD framework \cite{guinea2025} is the closest
antecedent, assembling elements from all three ICs---event-log discovery,
simulation, and ML prediction---but its evaluation remains component-wise,
placing it at L2 rather than a fully validated end-to-end pipeline. Crucially,
no paper in the reviewed corpus instantiates an \textbf{L3} architecture in
which process mining discovery feeds an absorbing Markov chain, whose transient
distributions drive Monte Carlo forecasting, with residuals corrected by a
learned model, all validated on SDLC event logs. PATHCAST is proposed as the
first such L3 system.

Quantitative evidence over the full 169-study set confirms the pattern. Only
1 of 169 papers (0.6\%) jointly satisfies IC1, IC2, and IC3---that is,
combines process mining, stochastic modeling, and forecasting in the same
study---and even that single paper does not specify a formal end-to-end
pipeline contract. Pairwise convergence is also low: IC1$\cap$IC2 (PM and
stochastic) appears in 10 papers (5.9\%), IC1$\cap$IC3 (PM and forecasting)
in 7 papers (4.1\%), and IC2$\cap$IC3 (stochastic and forecasting) in 18
papers (10.7\%). \textbf{No paper} in the reviewed corpus implements an
\textbf{L3} architecture. The literature has many partial bridges but no
end-to-end process-aware forecasting architecture for SDLC data. The
operational coding of the 73-PDF triangulation subset is consistent with the
full-set numbers and is summarized in
Figure~\ref{fig:pdf-integration-priority}.

\begin{figure}[p]
\centering
\begin{tikzpicture}
\begin{axis}[
  ybar,
  bar width=0.55cm,
  width=12cm, height=5.2cm,
  ylabel={Papers},
  symbolic x coords={L0,L1,L2,L3,High,Medium,Low},
  xtick=data,
  x tick label style={font=\small},
  ymin=0, ymax=60,
  ytick={0,10,20,30,40,50,60},
  nodes near coords,
  nodes near coords style={font=\small, above},
  fill=gray!45,
  draw=black,
  enlarge x limits=0.08,
]
\addplot coordinates {(L0,13) (L1,54) (L2,6) (L3,0) (High,54) (Medium,14) (Low,5)};
\end{axis}
\end{tikzpicture}
\caption{Operational integration levels and reading-priority distribution in the 73-paper PDF triangulation subset.}
\label{fig:pdf-integration-priority}
\end{figure}

\subsubsection{RQ3.2: Gap Analysis}

Three structural observations emerge from the 169-study extraction. First,
the field is \textbf{descriptively strong but forecastively weak}: 121
papers (71.6\%) discover or assess process behavior, but only 29 (17.2\%)
estimate future paths or probabilistic completion distributions. Second,
\textbf{evidence quality is uneven}: of the 169 confirmed studies, 69
(40.8\%) are case studies, 34 (20.1\%) are tool papers, 33 (19.5\%) are
theoretical, 13 (7.7\%) are simulation-only, 10 (5.9\%) are controlled
experiments, 5 (3.0\%) are surveys, and 5 (3.0\%) fall in residual
categories (mixed, framework, exploratory; sum: $69+34+33+13+10+5+5 = 169$).
The literature is therefore method-heavy and applied-evidence light. Third, \textbf{reproducibility is poor}: only 25 of 169 papers
(14.8\%) declare their dataset publicly, and only 4 (2.4\%) declare a
replication package. The QA scoring in Table~\ref{tab:qa-results} confirms
this pattern: only 33\% of studies score 1 on QA7 (data/code availability),
which is the single weakest QA criterion across the included set.

Taken together, these patterns explain why PATHCAST requires both a formal
pipeline and an evaluation strategy grounded in real SDLC traces. The gap is
not the absence of isolated techniques, but the absence of robust,
end-to-end, process-aware forecasting workflows validated on heterogeneous
software engineering data. The evidence-type distribution observed in the
73-PDF triangulation subset (Figure~\ref{fig:pdf-evidence-dist}) is
consistent with the full-set characterization above and is reported here as
an analytical view, not as a separate selection layer.

\begin{figure}[tbp]
\centering
\begin{tikzpicture}
\begin{axis}[
  xbar,
  bar width=0.45cm,
  width=12cm, height=5.5cm,
  xlabel={Papers},
  symbolic y coords={Method Proposal,Real-World Empirical,Experimental,Undefined,Computational Simulation},
  ytick=data,
  xmin=0, xmax=42,
  nodes near coords,
  nodes near coords align={horizontal},
  fill=gray!45,
  draw=black,
  enlarge y limits=0.18,
]
\addplot coordinates {
  (39,{Method Proposal})
  (13,{Real-World Empirical})
  (11,{Experimental})
  (7,{Undefined})
  (3,{Computational Simulation})
};
\end{axis}
\end{tikzpicture}
\caption{Evidence-type distribution in the 73-paper PDF triangulation subset.}
\label{fig:pdf-evidence-dist}
\end{figure}


\section{Discussion}
\label{sec:slr-discussion}

\subsection{Research Gap Identification}
\label{sec:slr-gaps}

The analysis of 2{,}340 screened records and 169 confirmed primary studies,
together with the criterion coverage patterns, supports five key findings:

\begin{description}

\item[F1: Descriptive Ceiling.]
121 of 169 confirmed papers (71.6\%) match IC1/IC4a (process mining in SE), but
only 29 (17.2\%) advance to IC3/IC4c (forecasting of process metrics). The
dominant contribution type is discovery or conformance checking, not
predictive analytics or distributional forecasting. This confirms the
descriptive ceiling reported anecdotally in prior surveys.

\item[F2: Event Log Construction Challenge.]
Across the 169 confirmed studies, only 25 papers (14.8\%) explicitly state
that their dataset is publicly available and only 4 papers (2.4\%) declare
any replication package. The majority of studies that mention data collection
procedures rely on ad hoc joins between version-control commits (84 papers,
49.7\%), issue trackers (37 papers, 21.9\%), IDE interaction logs (22 papers,
13.0\%), and CI/CD or build pipelines (15 papers, 8.9\%) rather than on a
reusable construction protocol with explicit case-correlation semantics. The
absence of standardized construction frameworks limits reproducibility and
cross-study comparability and is the most common cause of QA3 (data source
described reproducibly) failures observed in the QA scoring.

\item[F3: Stochastic Integration Gap.]
Only 52 of 169 confirmed papers (30.8\%) apply stochastic models (IC2/IC4b)
to software processes. The stochastic family is itself fragmented: 38 papers
(22.5\%) use Markov chains, 15 (8.9\%) use Monte Carlo simulation, 9 (5.3\%)
use stochastic Petri nets, and 8 (4.7\%) use system dynamics. Only 18 papers
(10.7\%) combine stochastic modeling with forecasting (IC2 $\cap$ IC3), and
only 10 papers (5.9\%) combine stochastic modeling with process mining (IC1
$\cap$ IC2). Markov chains and Monte Carlo simulation are used extensively in
BPM and software reliability engineering, but their integration with process
mining for SDLC forecasting is minimal. EC3 being the second-largest
exclusion reason at FT ($n = 197$, 27.5\% of 717 FT excludes) reinforces the
point: the stochastic modeling community has not converged on applying its
techniques to software development process data.

\item[F4: Pipeline Fragmentation.]
Studies apply techniques in isolation. The formal composition of a
PM $\rightarrow$ Markov $\rightarrow$ MC $\rightarrow$ ML pipeline with
inter-stage data contracts has not been proposed in the literature for SDLC
contexts. The IC-combination analysis on the 169-study set is unequivocal:
the joint convergence of process mining, stochastic modeling, and forecasting
(IC1 $\cap$ IC2 $\cap$ IC3) appears in only 1 paper (0.6\%), and even that
single paper does not specify a formal end-to-end pipeline contract.
Pairwise convergences are also low (IC1 $\cap$ IC2: 10 papers, 5.9\%;
IC1 $\cap$ IC3: 7 papers, 4.1\%; IC2 $\cap$ IC3: 18 papers, 10.7\%). The
closest antecedent in the corpus is PRIMAD \cite{guinea2025}, which assembles
discovery, simulation, and learning components but validates each
independently and leaves the unified pipeline unevaluated. Even when
stochastic components are present, validation usually stops at process
analysis or simulation without a downstream learning component.

\item[F5: Process-Derived Features Unexplored.]
ML-based prediction studies in the confirmed set rely overwhelmingly on
traditional features (code metrics, commit characteristics, defect counts)
rather than on features derived from process mining. Among the 32 studies in
the 169-set whose research contribution centers on machine-learning
prediction, only 1 study explicitly couples ML with a named process-mining
algorithm (alpha, inductive, heuristic, conformance, fuzzy, or
Petri-net-based discovery). Conformance scores, rework counts,
Markov-expected remaining time, and path entropy---all features that are
naturally produced by a process-mining stage---remain essentially absent
from the SE prediction literature.

\end{description}

These five findings map directly to the five deficits identified in
Section~\ref{sec:gap-analysis}: F1 corresponds to D1 (Descriptive Ceiling),
F3 to D3 (Stochastic Integration Absence), F4 to D5 (Pipeline Fragmentation),
and F5 to D4 (Feature Engineering Void). F2 provides empirical evidence for a
challenge that PATHCAST addresses through Stage~1 (Event Log Construction).


\subsection{Positioning of PATHCAST}
\label{sec:slr-positioning}

Table~\ref{tab:positioning} positions PATHCAST against the integration levels
identified in the 169-study confirmed set, using both the IC-combination
counts derived from the structured extraction and the operational coding
obtained from the 73-PDF triangulation subset.

\begin{table}[htbp]
\centering
\caption{Integration levels in the 169-study confirmed set and PATHCAST positioning.}
\label{tab:positioning}
\small
\setlength{\tabcolsep}{4pt}
\begin{tabular}{c p{4.8cm} c c p{3.4cm}}
\toprule
\textbf{Level} & \textbf{Description} & \textbf{Papers (169-set, IC-based)}
  & \textbf{Papers (73-PDF subset, operational)} & \textbf{Representative example} \\
\midrule
L0 & Single analytical family ($\leq$1 IC matched); no cross-technique chaining
  & 95 & 13 & Contextual or single-technique studies \\
L1 & Two ICs matched; two families present, but no explicit reproducible pipeline contract
  & 70 & 54 & Stochastic models or PM without formal coupling \\
L2 & Three ICs matched; partial bridge linking PM, stochastic, and forecasting, but incomplete pipeline or missing ML closure
  & 4 & 6 & PRIMAD \cite{guinea2025}; Incerto \cite{incerto2025}; L\'opez-Pintado \cite{lopezpintado2023} \\
L3 & Four ICs matched; unified framework: PM $\rightarrow$ Markov $\rightarrow$ MC $\rightarrow$ ML with formal inter-stage contracts
  & 0 & 0 & \textbf{PATHCAST} (proposed) \\
\midrule
\textbf{Total} & --- & \textbf{169} & \textbf{73} & --- \\
\bottomrule
\end{tabular}
\end{table}

PATHCAST is positioned at level L3---the first proposal, to the best of the
author's knowledge, to formally compose process mining, absorbing Markov chain
analysis, Monte Carlo forecasting, and ML enhancement into a single
architecturally specified pipeline applied to SDLC data.


\subsection{Proposed Taxonomy: SPMF}
\label{sec:slr-taxonomy}

Based on the systematic analysis, this review proposes the Software Process
Mining and Forecasting (SPMF) taxonomy with three dimensions:

\begin{description}
\item[Analytical Depth] classifies studies along four levels: L1
  (Descriptive: discovery, conformance), L2 (Diagnostic: bottleneck, rework),
  L3 (Predictive: outcome, remaining time), and L4 (Forecasting:
  distributional, probabilistic).

\item[SDLC Coverage] maps which lifecycle phases are addressed: Requirements
  and Design, Development (coding, commits, PRs), Quality (testing, reviews,
  bugs), Delivery (CI/CD, release), and Operations (incidents, maintenance).

\item[Technique Integration] classifies methodological combination:
  Single (PM only, ML only, MC only), Dual (PM + ML, PM + Markov, Markov + MC),
  and Pipeline (PM $\rightarrow$ Markov $\rightarrow$ MC $\rightarrow$ ML).
\end{description}


\subsection{Research Agenda}
\label{sec:slr-agenda}

The gaps identified motivate four research directions:

\begin{description}
\item[RA1: Standardized Event Log Construction for SDLC.]
  Frameworks for transforming heterogeneous repository data into comparable
  event logs with cross-tool case correlation. Maps to Stage~1 of PATHCAST.

\item[RA2: Process-Aware Stochastic Forecasting.]
  Formal integration of process mining with Markov chains and Monte Carlo
  simulation for distributional SDLC forecasting. Maps to Stages~2--4 of
  PATHCAST.

\item[RA3: Process-Derived Features for ML.]
  Systematic exploration of features derived from process mining for outcome
  prediction. Maps to the ML Enhancement Component of PATHCAST.

\item[RA4: Empirical Benchmarking.]
  Evaluation of forecasting methods on comparable OSS repository datasets with
  temporal hold-out and standardized baselines. Maps to the empirical
  evaluation in Chapter~\ref{ch:evaluation}.
\end{description}


\subsection{Implications for Practitioners}
\label{sec:slr-practitioners}

Practitioners developing software measurement programs can extract three
actionable insights from the 169-study confirmed set. First, the availability
of structured event log data in issue tracking systems (Jira, GitHub Issues)
is a prerequisite for process mining; organizations should invest in
consistent workflow state management and transition logging before attempting
process analysis. Second, open-source toolkits (PM4Py, ProM, Disco) suffice
for discovery and conformance checking in most SDLC contexts, as suggested by
the dominance of the inductive miner (51 papers) and alpha miner (45 papers)
in the included studies' tool usage patterns. Third, the stochastic extension
(Markov chains for transition probability estimation, Monte Carlo for
throughput forecasting) requires less infrastructure than typically assumed:
the data already exists in the event logs; the challenge is methodological
composition, not data collection.


\section{Threats to Validity of the SLR}
\label{sec:slr-threats}

\subsection{Construct Validity}

The search strings may exclude studies that use non-canonical terminology.
This threat is mitigated through three mechanisms: the complementary MSR
string, the frontier stochastic-PM string, and snowballing. The 10/10 control
paper recovery provides empirical evidence of high recall.

\subsection{Internal Validity}

To further strengthen the reliability of the screening process, a human spot-check of rating disagreements is planned. A sample of approximately 50 papers where the two primary reviewers disagreed on the final decision (include/exclude) will be analyzed by a third senior researcher to identify the root cause of the disagreement (e.g., criteria ambiguity, reviewer error). The findings from this analysis will be incorporated into the camera-ready version of this work to provide a qualitative assessment of reviewer alignment.

LLM-based screening introduces the risk of systematic decision bias. This
threat is addressed by four mechanisms.

(i)~\textbf{Structured outputs}. All screening responses use a strict JSON
schema, enabling post-hoc analysis of decision patterns and rationale
distributions.

(ii)~\textbf{Manual verification of all \texttt{include} decisions}.
Every paper retained at the FT stage was inspected by the author against the
PRISMA flow before being added to the 169-study confirmed set.

(iii)~\textbf{Cross-model inter-rater agreement (Cohen's $\kappa$)}.
A stratified random 20\% sample of T/A ($n_{\text{theoretical}} = 468$ from
the 2{,}340-paper working set) and FT ($n_{\text{theoretical}} = 177$ from
the 886-paper FT queue) decisions was independently re-rated by a different
model family (\textit{claude-sonnet-4-6}) acting as a second rater. Of the
468 T/A papers in the sample, 420 produced a parseable decision from the
verifier and entered the agreement computation; the remaining 48 papers
returned a non-parseable response (typically because the abstract was
absent, leading the verifier to refuse a categorical answer) and were
excluded from the $\kappa$ calculation as missing-at-random. All 177 FT
sample papers produced a parseable decision. Both the multi-class
$\kappa$ (over the original \texttt{include}/\texttt{exclude}/\texttt{maybe}
or \texttt{include}/\texttt{exclude}/\texttt{pending} categories) and the
binary $\kappa$ (collapsing the rater outputs to
\texttt{include} versus \texttt{not-include}, the only distinction that
affects the final selection set) are reported in
Table~\ref{tab:kappa-results}. The binary $\kappa$ is the operationally
relevant measure for SLR selection outcomes. Both binary $\kappa$ values
($0.695$ for T/A and $0.694$ for FT) exceed the substantial-agreement
threshold $\kappa \geq 0.61$ \cite{landiskoch1977}. The multi-class FT
$\kappa$ is lower ($0.479$, moderate) because the verifier returned
\texttt{pending} on $25/177$ FT papers where the primary screener committed
to a binary decision; this represents a difference in conservativeness, not
a difference in the final selection outcome. The cross-model verification
design follows the guidance of Khraisha~et~al.~\cite{khraisha2024} and
Syriani~et~al.~\cite{syriani2023} for LLM-assisted SLRs and is intended as a
tractable proxy for a fully independent human second-rater protocol; a human
spot-check on disagreement cases is recommended before camera-ready
submission.

\input{results/kappa/kappa_report}

(iv)~\textbf{T/A false-positive rate}.
The LLM confirmatory pass at FT level (Round~1) rejected 33 of 108 T/A
\texttt{include} papers (30.6\%), primarily for EC3 (algorithm without SE
evaluation). This false-positive rate is consistent with the
T/A-screening false-positive rates reported by Khraisha~et~al.\
\cite{khraisha2024} for medical SLRs (in the 25--35\% range, depending on
domain specificity) and by Syriani~et~al.~\cite{syriani2023} for SE
SLRs. False-positives are caught at the FT phase before any synthesis claim
is made and therefore do not affect the F1--F5 evidence base. The Round~2
screening reprocessed 495 \texttt{maybe} papers with enriched abstracts and
brought the confirmed set to its final value of 169.

\subsection{External Validity}

The review is limited to English-language peer-reviewed publications,
potentially excluding relevant work in other languages or grey literature.
The broad time span (1994--2026) and five-database coverage mitigate
geographic and temporal bias. The deferred high-recall auxiliary corpus
(3{,}441 records from SpringerLink, WoS, Snowball) provides a mechanism for
expansion if gaps in the confirmed 169-study set are identified.

\subsection{Conclusion Validity}

Synthesis findings F1--F5 are derived from the structured data extraction
applied to the full 169-study confirmed set
(Section~\ref{sec:slr-extraction}; Table~\ref{tab:extraction}), with the
73-PDF subset used as a triangulation layer rather than as the primary
evidence base. The IC-combination counts and the technique distributions
that ground F1, F3, and F4 are computed directly from the extraction CSV;
the QA-based filtering reported in Table~\ref{tab:qa-results} retains 136 of
169 studies (80.5\%) for the synthesis and the remaining 33 are excluded
from the F1--F5 evidence base.

A residual conclusion-validity threat is the LLM-based extraction of
PM-technique, stochastic-technique, and SDLC-phase fields. To mitigate this,
the 73-PDF triangulation subset was used to spot-check the extracted fields
against the corresponding full texts; no systematic disagreement was
observed in the spot-check, but a complete human re-extraction on a
stratified sample is recommended before camera-ready submission.


\subsection{Reproducibility and Use of AI}
\label{sec:slr-replication}

In line with current Elsevier and Springer-Nature editorial policy, this
review explicitly declares the use of AI tools in its conduct. The primary
screener at the title/abstract and full-text stages is
\textit{claude-haiku-4-5-20251001} (Anthropic), invoked through the
Anthropic Message Batches API for screening and through the
synchronous Messages API for QA scoring. The cross-model
inter-rater verification described above uses
\textit{claude-sonnet-4-6} (Anthropic) as the second rater. AI tools are
\textbf{not} listed as authors of this work and are responsible for none of
the editorial judgements; final selection, QA threshold, taxonomy design,
and synthesis claims are the responsibility of the author.

A complete replication package is deposited at Zenodo (DOI~10.5281/zenodo.15719919)
and includes: the search strings for all five databases, the deduplication and
enrichment scripts (\texttt{pipeline/dedup.py}, \texttt{pipeline/enrich.py}),
the screening prompts and JSON schema (\texttt{config/screening\_criteria.py}),
the QA prompt (\texttt{pipeline/qa\_llm.py}), the cross-model verification
prompts and computation (\texttt{pipeline/kappa.py}), the decision CSVs at
each stage (\texttt{results/screening/ta\_screening\_results.csv},
\texttt{results/screening/ft\_screening\_results.csv}), the QA-scored 169-study
table (\texttt{results/qa\_assessment.csv}), the structured extraction
table (\texttt{results/extraction/extraction\_template.csv}), the
inter-rater agreement reports (\texttt{results/kappa/}), and a PRISMA flow
snapshot. The package is intended to be self-contained: any reviewer can
re-run the LLM-assisted protocol on the same working set and reproduce the
169-study confirmed selection within the stated $\kappa$ confidence band.


\section{Chapter Summary}
\label{sec:slr-summary}

This chapter presented a systematic literature review of the intersection of
process mining, stochastic modeling, and machine learning applied to software
process forecasting. The review executed three complementary search strings
across five databases, yielding 8{,}347 records, 5{,}783 after deduplication,
and a 2{,}340-paper operational working set for systematic screening. Title and
abstract screening using a LLM-assisted protocol produced 111 candidate
\texttt{include} decisions and 775 \texttt{maybe} decisions (886 papers for
full-text review).

Full-text screening combined LLM-assisted passes (Rounds~1 and~2, with cascade
abstract enrichment raising coverage from 16.6\% to 72.6\%, i.e., 643/886
papers) with a final \emph{screen-blanks} pass that processed the 186 papers
that still lacked an FT decision after the earlier rounds. In the final
screening artifact, 238/886 papers still lack an abstract field, but all have
a recorded FT decision. Papers whose full text remained inaccessible despite
all retrieval attempts were closed under EC5 (148 papers). The final result is
\textbf{169 confirmed primary studies} and 717 FT excluded (EC3: 229, EC1: 205,
EC5: 148, EC2: 110, EC4: 60), with zero pending decisions. Data extraction
for all 169 studies was completed using a structured LLM-assisted pipeline
operating on locally available PDFs (75/169, with 73 operationally coded in
the triangulation subset) or abstract/title metadata (94/169).

Analysis of the 169 confirmed studies supports five key findings. First, the
field exhibits a descriptive ceiling: 71.6\% of confirmed papers apply process
mining to software artifacts (IC1), but only 17.2\% advance to forecasting of
process metrics (IC3)~(F1). Second, event log construction from heterogeneous
repository sources is a pervasive challenge, with ad hoc joins between issues,
commits, and pipeline logs the norm~(F2). Third, the integration of stochastic
models with process mining for SDLC forecasting is minimal---only 10.7\% of
confirmed papers combine IC2 and IC3~(F3). Fourth, no confirmed study implements
a formally composed PM $\rightarrow$ Markov $\rightarrow$ MC $\rightarrow$ ML
pipeline for software processes; PRIMAD~\cite{guinea2025} is the closest
antecedent at L2~(F4). Fifth, ML studies in SE rarely incorporate process-derived
features~(F5).

Synthesis of the top-ranked papers by IC cluster sharpens the picture. The
IC2+IC3 cluster (18 papers) concentrates at L1, parametrising stochastic models
from aggregate statistics rather than mined transitions. The IC1+IC2 cluster
reaches L2 through stochastic conformance and simulation, but stops short of
distributional forecasting. The IC1+IC3 cluster chains PM and prediction
sequentially without a stochastic layer. The IC2+IC4 cluster demonstrates that
public repository traces support state-transition modelling, but event-to-state
mapping remains unstandardised. No paper occupies L3.

These five gaps directly motivate the PATHCAST method presented in
Chapter~\ref{ch:method}. PATHCAST is positioned at integration level~L3 in
the SPMF taxonomy---the first formally specified pipeline to compose all four
technique families into a unified, architecturally specified framework for
SDLC process forecasting.

Methodologically, the LLM-assisted protocol used in this review achieves
substantial inter-rater agreement under cross-model verification (binary
$\kappa = 0.695$ for T/A and $\kappa = 0.694$ for FT) and full QA coverage of
the 169 confirmed studies (mean QA score $= 4.88/8$, with 136/169 retained
above the $\geq 4/8$ threshold). The complete replication package and the
inter-rater agreement reports are deposited at Zenodo
(DOI~10.5281/zenodo.15719919).
