%% =============================================================================
%% IST submission — main entry point
%%
%% Build:
%%   pdflatex main
%%   bibtex main
%%   pdflatex main && pdflatex main
%%
%% Required \input files (preserve relative paths under the project root):
%%   ../results/qa_assessment_summary.tex
%%   ../results/kappa/kappa_report.tex
%%   ../results/auxiliary/aux_qa_summary.tex
%%   ../results/auxiliary/aux_ft_summary.tex
%%   ../results/auxiliary/aux_reft_summary.tex
%%   ../results/auxiliary/kappa/aux_kappa_report.tex
%%   ../results/ec5_recovery/ec5_recovery_report.tex
%%   ../results/sensitivity/sensitivity_report.tex
%% Bibliography:
%%   ../results/final_review/missing_references.bib (+ thesis bib if needed)
%% =============================================================================

\documentclass[review,3p,times]{elsarticle}

%% --- Math + tables ---
\usepackage{amsmath,amssymb}
\usepackage{booktabs}
\usepackage{multirow}
\usepackage{array}
\usepackage{tabularx}
\usepackage{makecell}
\usepackage{longtable}
\usepackage{caption}

%% --- Figures ---
\usepackage{graphicx}
\usepackage{tikz}
\usetikzlibrary{arrows.meta,positioning,shapes}
\usepackage{pgfplots}
\pgfplotsset{compat=1.18}

%% --- Misc ---
\usepackage{enumitem}
\usepackage{xcolor}
\usepackage{hyperref}
\hypersetup{colorlinks=true, linkcolor=blue!50!black, citecolor=blue!50!black, urlcolor=blue!50!black}

%% --- File-search path so \input picks up results/ from project root ---
\makeatletter
\providecommand*\input@path{}
\g@addto@macro\input@path{{../}}
\makeatother

%% --- Bibliography style: Elsevier numeric + sorted ---
\bibliographystyle{elsarticle-num}

\journal{Information and Software Technology}

\begin{document}

%% =========================
%% Front matter
%% =========================
\begin{frontmatter}

\title{Process Mining and Stochastic Modeling for Software Process
       Forecasting: A Systematic Literature Review with an LLM-Assisted Protocol}

%% Authors
\author[pucpr]{Rodrigo Almeida de Oliveira\corref{cor1}\fnref{orcid1}}
\ead{rodrigo1.almeida@pucpr.edu.br}

\author[pucpr]{Juliano de Paulo Ribeiro\fnref{orcid2}}
\ead{juliano.ribeiro@pucpr.edu.br}

\author[pucpr]{Edson Emilio Scalabrin\fnref{orcid3}}
\ead{scalabrin@ppgia.pucpr.br}

\cortext[cor1]{Corresponding author.}

\fntext[orcid1]{ORCID: 0009-0009-6310-4126}
\fntext[orcid2]{ORCID: 0009-0004-3605-485X}
\fntext[orcid3]{ORCID: 0000-0002-3918-1799}

\affiliation[pucpr]{
  organization={Pontif\'icia Universidade Cat\'olica do Paran\'a (PUCPR), Graduate Program in Informatics (PPGIa)},
  city={Curitiba},
  state={PR},
  country={Brazil}
}

%% --- Abstract (150–300 words, structured) ---
\begin{abstract}
\textbf{Context.}
Software development processes generate rich event data --- commits, issues,
pull requests, CI/CD logs --- that process mining (PM) and stochastic models
can analyze to support forecasting. Yet integration of these techniques for
software development lifecycle (SDLC) prediction is reported only
anecdotally.
\textbf{Objective.}
To map the state of the art at the intersection of process mining,
stochastic modeling, and machine-learning forecasting applied to software
processes; to identify research gaps; and to propose a research agenda.
\textbf{Method.}
We executed a systematic literature review following Kitchenham and Charters
(2007), Petersen et~al. (2015) for mapping, and Wohlin (2014) for
snowballing, augmented with an LLM-assisted screening protocol
(\textit{claude-haiku-4-5} as primary screener, \textit{claude-sonnet-4-6}
as cross-model verifier). The protocol screened 5{,}783 deduplicated records
across two tiers: a 2{,}340-paper working-set tier and a 3{,}807-paper
auxiliary tier. Cohen's $\kappa$ inter-rater agreement was computed on
stratified 20\% samples; 8-criterion quality assessment (Dyba and Dingsoyr,
2008) was applied to all confirmed includes.
\textbf{Results.}
404 primary studies were confirmed (169 working-set + 212 first-pass auxiliary + 23 second-pass auxiliary after abstract enrichment).
Inter-rater agreement was substantial in the working-set tier
($\kappa_{\text{binary}} = 0.695$ T/A, $0.694$ FT). Five findings emerge:
F1 a descriptive ceiling (PM dominant, forecasting minority); F2 ad-hoc
event-log construction; F3 weak stochastic--PM integration; F4 pipeline
fragmentation (only 1 of 404 papers jointly satisfies PM + stochastic +
forecasting); F5 process-derived ML features unexplored.
\textbf{Conclusion.}
The literature is descriptively rich but predictively fragmented. We
propose the SPMF taxonomy and a four-direction research agenda; PATHCAST is
introduced as the first formally specified L3 pipeline for SDLC
forecasting.
\end{abstract}

\begin{keyword}
Systematic Literature Review \sep
Process Mining \sep
Stochastic Modeling \sep
Software Process Forecasting \sep
LLM-Assisted Screening \sep
Mining Software Repositories
\end{keyword}

%% --- Highlights (3–5 bullets, ≤85 chars each) ---
\begin{highlights}
\item LLM-assisted SLR protocol screens 5{,}783 deduplicated software-process papers in two corpus tiers.
\item Cross-model Cohen's $\kappa = 0.695$ inter-rater agreement on a 20\% stratified sample.
\item 404 primary studies confirmed; only 1 jointly satisfies PM, stochastic, and forecasting.
\item Five findings (F1--F5) and SPMF taxonomy proposed; four-direction research agenda.
\item Full replication package (prompts, scripts, decisions) deposited at Zenodo.
\end{highlights}

\end{frontmatter}

%% =========================
%% Body
%% =========================
\input{cap3_article_body.tex}

%% =========================
%% Back matter — declarations
%% =========================
\section*{CRediT authorship contribution statement}
\textbf{Rodrigo Almeida de Oliveira:} Conceptualization, Methodology,
Software, Investigation, Data curation, Formal analysis, Visualization,
Writing -- original draft, Writing -- review \& editing.
\textbf{Juliano de Paulo Ribeiro:} Methodology, Validation, Investigation,
Writing -- review \& editing.
\textbf{Edson Emilio Scalabrin:} Conceptualization, Methodology,
Supervision, Writing -- review \& editing, Resources, Funding acquisition.

\section*{Declaration of competing interest}
The authors declare that they have no known competing financial interests
or personal relationships that could have appeared to influence the work
reported in this article.

\section*{Funding}
This study was financed in part by the Coordena\c c\~ao de Aperfei\c
coamento de Pessoal de N\'ivel Superior --- Brasil (CAPES) --- Finance
Code 001, through a doctoral scholarship granted to the corresponding
author within the Graduate Program in Informatics at the Pontif\'icia
Universidade Cat\'olica do Paran\'a (PPGIa/PUC-PR).

\section*{Declaration of generative AI in the writing process}
During the conduct of this systematic literature review, the authors used
\textit{claude-haiku-4-5-20251001} (Anthropic) as the primary LLM screener
at the title/abstract and full-text stages and as the LLM scorer for the
quality-assessment rubric, and \textit{claude-sonnet-4-6} (Anthropic) as
the cross-model verifier for inter-rater agreement. All decisions were
inspected by the authors; the LLM tools are not listed as authors and
bear no responsibility for editorial judgements. Full prompts, JSON
schemas, and per-paper decisions are deposited in the replication package
(\href{https://doi.org/10.5281/zenodo.15719919}{Zenodo
DOI~10.5281/zenodo.15719919}). The authors reviewed and edited the content
as needed and take full responsibility for the content of the
publication.

\section*{Data availability}
A full replication package is available at Zenodo
(\href{https://doi.org/10.5281/zenodo.15719919}{DOI~10.5281/zenodo.15719919}).
The package contains: search strings for all five databases, deduplication
and enrichment scripts, screening prompts and JSON schemas, the QA prompt,
the cross-model $\kappa$ verification scripts, all decision CSVs at every
pipeline stage, the QA-scored 169-study and 212-study tables, the
structured extraction tables (working-set and auxiliary), the inter-rater
agreement reports, and a PRISMA flow snapshot.

%% --- Bibliography ---
\bibliography{../results/final_review/missing_references}

\end{document}
